\begin{optDefinition}
\bibannex
\begin{references}

\reference
    {Alberga, C.N., Bosman-Clark, C., Mikelsons, M., Van Deusen, M.,
     \& Padget, J.A.,}
    {Experience with an Uncommon LISP,}
    {Proceedings of 1986 ACM Symposium on LISP and Functional Programming, ACM,
     New York, 1986 (also available as IBM Research Report RC-11888).}
    \label{lisp/vm}

\reference
    {Berrington N., Deroure D. \& Padget J.A.,}
    {Guaranteeing Unpredictability,}
    {in preparation.}
    \label{lisp-processes}

\reference
    {Bobrow D.G., DiMichiel L.G., Gabriel R.P., Keene S.E, Kiczales G.
     \& Moon D.A,}
    {Common Lisp Object System Specification,}
    {SIGPLAN Notices, Vol. 23, September 1988.}
    \label{clos}

\reference
    {Chailloux J., Devin M. \& Hullot J-M.,}
    {LELISP: A Portable and Efficient Lisp System,}
    {Proceedings of 1984 ACM Symposium on Lisp and Functional Programming,
     Austin, Texas, pp113-122, published by ACM Press, New York.}
    \label{le-lisp}

\reference
    {Chailloux J., Devin M., Dupont F., Hullot J-M., Serpette B.,
     \& Vuillemin J.,}
    {Le-Lisp de l'INRIA, Version 15.2, Manuel de r\'ef\'erence,}
    {INRIA, Rocquencourt, May 1987.}
\label{lelisp-manual}

\reference
    {Clinger W. \& Rees J.A. (eds.),}
    {The Revised\/$^3$ Report on Scheme,}
    {SIGPLAN Notices, Vol. 21, No. 12, 1986.}
    \label{scheme-3}

\reference
    {Cointe P.,}
    {Metaclasses are First Class: the ObjVlisp model,}
    {Proceedings of OOPSLA '87, published as SIGPLAN Notices, Vol 22,
     No 12 pp156-167.}
    \label{objvlisp}

\reference
    {Fitch J.P. \& Norman A.C.,}
    {Implementing Lisp in a High-Level Language,}
    {Software Practice and Experience, Vol 7, pp713-725.}
    \label{cambridge-lisp}

\reference
    {Friedman D. \& Haynes C.,}
    {Constraining Control,}
    {Proceedings of 11th Annual ACM
     Symposium on Principles of Programming Languages, pp245-254,
     published by ACM Press, New York, 1985.}
    \label{constraining}

\reference
    {Hudak P. \& Wadler P., (eds.)}
    {Report on the Functional Programming Language Haskell,}
    {Yale University, Department of Computer Science, Research
     Report YALEU/DCS/RR-666, December 1988.}
    \label{haskell}

\reference
    {Landin P.J.,}
    {The Next 700 Programming Languages,}
    {Communications of the ACM, Vol 9, No 3., 1966, pp156-166.}
    \label{iswim}

\reference
    {Lang K.J. \& Pearlmutter B.A.,}
    {Oaklisp: An Object-Oriented Dialect of Scheme,}
    {Lisp and Symbolic Computation, Vol. 1, No. 1, June 1988, pp39-51,
     published by Kluwer Academic Publishers, Boston.}
    \label{oaklisp}

\reference
    {MacQueen D., {\em et al,}}
    {Modules for Standard ML,}
    {Proceedings of 1984 ACM Symposium on Lisp and Functional Programming,
     Austin, Texas, pp198-207, published by ACM Press, New York.}
    \label{ML-modules}

\reference
    {Milner R., {\em et al,}}
    {Standard ML,}
    {Laboratory for the Foundations of Computer Science,
     University of Edinburgh, Technical Report.}
    \label{std-ml}

\reference
    {Padget J.A., {\em et al,}}
    {Desiderata for the Standardisation of Lisp,}
    {Proceedings of 1986 ACM Conference on Lisp and Functional Programming,
     pp54-66, published by ACM Press, New York, 1986.}
    \label{desiderata}

\reference
    {Pitman K.M.,}
    {An Error System for Common Lisp,}
    {ISO WG16 paper N24.}
    \label{Pit-errors}

\reference
    {Rees J.A.,}
    {The T Manual,}
    {Yale University Technical Report, 1986.}
    \label{t-manual}

\reference
    {Slade S.,}
    {The T Programming Language, a Dialect of Lisp,}
    {Prentice-Hall 1987.}
    \label{t-book}

\reference
    {Shalit A.,}
    {Dylan, an object-oriented dynamic language,}
    {Apple Computer Inc., 1992.}
    \label{dylan}

\reference
    {Steele G.L. Jr.,}
    {Common Lisp the Language,}
    {Digital Press, 1984.}
    \label{CLtl}

\reference
    {Steele G.L. Jr.,}
    {Common Lisp the Language (second edition),}
    {Digital Press, 1990.}
    \label{CLtlt}

\reference
    {Stoyan H. {\em et al,}}
    {Towards a Lisp Standard,}
    {published in the Proceedings of the 1986 European Conference on Artificial
     Intelligence.}
    \label{stoyan}

\reference
    {Teitelman W.,}
    {The Interlisp Reference Manual,}
    {Xerox Palo Alto Research Center, 1978.}
    \label{interlisp}

\reference
    {Bretthauer, H. and Kopp, J.,}
    {The Meta-Class-System MCS. A Portable Object System for Common Lisp.
     -Documentation-.}
    {Arbeitspapiere der GMD 554, Gesellschaft fur Mathematik und
     Datenverarbeitung (GMD), Sankt Augustin (July 1991).}
    \label{mcs}

\end{references}
\end{optDefinition}
