\sclause{Copying}
\label{copy}
\index{general}{level-0 modules!copy}
\index{general}{copy!module}
%
\begin{optPrivate}
    I have a cryptic note in my hardcopy which says {\tt (find-method copy
        integer)}, but I don't know why.
\end{optPrivate}
%
\begin{optDefinition}
\noindent
The defined name of this module is {\tt copy}.

\generic{deep-copy}
%
\begin{genericargs}
    \item[object] An object to be copied.
\end{genericargs}
%
\result%
Constructs and returns a copy of the source which is the same (under some class
specific predicate) as the source and whose slots contain copies of the objects
stored in the corresponding slots of the source, and so on.  The exact behaviour
for each class of {\em object\/} is defined by the most applicable method for
{\em object}.
%
\seealso%
Class specific sections which define methods on \genericref{deep-copy}:
list~(\ref{list}), string~(\ref{string}), table~(\ref{table}) and
vector~(\ref{vector}).

\method{deep-copy}{object}
%
\begin{specargs}
    \item[object, \classref{object}] An object.
\end{specargs}
%
\result%
Returns {\em object}.

\method{deep-copy}{class}
%
\begin{specargs}
    \item[class, \classref{class}] A class.
\end{specargs}
%
\result%
Constructs and returns a new structure whose slots are initialized with copies
(using \genericref{deep-copy}) of the contents of the slots of {\em class}.

\generic{shallow-copy}
%
\begin{genericargs}
    \item[object] An object to be copied.
\end{genericargs}
%
\result%
Constructs and returns a copy of the source which is the same (under
some class specific predicate) as the source.  The exact behaviour for
each class of {\em object\/} is defined by the most applicable method
for {\em object}.
%
\seealso%
Class specific sections which define methods on \genericref{shallow-copy}:
pair~(\ref{pair}), string~(\ref{string}), table~(\ref{table}) and
vector~(\ref{vector}).

\method{shallow-copy}{object}
%
\begin{specargs}
    \item[object, \classref{object}] An object.
\end{specargs}
%
\result%
Returns {\em object}.
%
\method{shallow-copy}{class}
%
\begin{specargs}
    \item[class, \classref{class}] A class.
\end{specargs}
%
\result
Constructs and returns a new structure whose slots are initialized
with the contents of the correpsonding slots of {\em struct}.
%
\end{optDefinition}
