\sclause{Floating Point Numbers}
\index{general}{float}
\index{general}{float!module}
\label{float}
\index{general}{level-0 modules!float}
%
\begin{optDefinition}
\noindent
The defined name of this module is {\tt float}.  This module defines the
abstract class \classref{float}\ and the behaviour of some generic functions on
floating point numbers.  Further operations on numbers are defined in the
numbers module (\ref{number}) and further operations on floating point numbers
are defined in the elementary functions module (\ref{elementary-functions}).  A
concrete float class is defined in the double float module (\ref{double-float}).

\syntaxform{float}
%
The syntax for the external representation of floating point literals is defined
in syntax table~\ref{float-syntax}.  The representation used by
\functionref{write} and \functionref{prin} is that of a sign, a whole part and a
fractional part without an exponent, namely that defined by {\em float format
    3}.  Finer control over the format of the output of floating point numbers
is provided by some of the formatting specifications of \functionref{format}
(section~\ref{formatted-io}).
%
\Syntax
\label{float-syntax}
\defSyntax{float}{
\begin{syntax}
    \scdef{float}: \\
    \>  \scoptref{sign} \scref{unsigned-float} \scoptref{exponent} \\
   \scdef{unsigned-float}: \\
   \>  \scref{float-format-1} \\
   \>  \scref{float-format-2} \\
   \>  \scref{float-format-3} \\
   \scdef{float-format-1}: \\
   \>  \scref{decimal-integer} . \\
   \scdef{float-format-2}: \\
   \>  . \scref{decimal-integer} \\
   \scdef{float-format-3}: \\
   \>  \scref{float-format-1} \scref{decimal-integer} \\
   \scdef{exponent}: \\
   \>  \scref{double-exponent}
\end{syntax}}%
\showSyntaxBox{float}

A floating point number\index{general}{external representation!floating point}
has six forms of external representation depending on whether either or both the
whole and the fractional part are specified and on whether an exponent is
specified.  In addition, a positive floating point number is optionally preceded
by a plus sign and a negative floating point number is preceded by a minus sign.
For example:
%
\verb|+123.| ({\em float format 1\/}),
\verb|-.456| ({\em float format 2\/}),
\verb|123.456| ({\em float format 3\/}); and with exponents:
\verb|+123456.D-3|,
\verb|1.23455D2|,
\verb|-.123456D3|.

\derivedclass{float}{number}
\index{general}{level-0 classes!\theclass{float}}
%
The abstract class which is the superclass of all floating point
numbers.
%
\function{floatp}
\begin{arguments}
    \item[objext] Object to examine.
\end{arguments}
%
\result%
Returns {\em object\/} if it is a floating point number, otherwise \nil{}.

\generic{ceiling}
%
\begin{genericargs}
    \item[float, \classref{float}] A floating point number.
\end{genericargs}
%
\result%
Returns the smallest integral value not less than {\em float\/} expressed as a
float of the same class as the argument.

\generic{floor}
%
\begin{genericargs}
    \item[float, \classref{float}] A floating point number.
\end{genericargs}
%
\result%
Returns the largest integral value not greater than {\em float\/} expressed as a
float of the same class as the argument.

\generic{round}
%
\begin{arguments}
    \item[float] A floating point number.
\end{arguments}
%
\result%
Returns the integer whose value is closest to {\em float}, except in the case
when {\em float\/} is exactly half-way between two integers, when it is rounded
to the one that is even.

\generic{truncate}
%
\begin{arguments}
    \item[float] A floating point number.
\end{arguments}
%
\result%
Returns the greatest integer value whose magnitude is less than or equal to {\em
    float}.
%
\end{optDefinition}
