\begin{introduction}
\label{sec:intro}
\begin{optPrivate}
    The behaviour of syntax operators is never explained, especially with
    respect to how many times they might be expanded.

    The hierarchy of types (classes) is not specified anywhere.  The coercion
    ladder, if taken as part of the hierarchy, is upside down in some places (if
    the class systems is taken as prototypical of types).

    Nothing about the behaviour of \telos\ is specified.  How is the CPL
    calculated?  When are methods invoked?  What is method combination?  (Note:
    it seems someone else wrote this part.  It seems much less smooth than the
    rest).

    Threads are hardly discussed.  How can tail-recursion be specified
    linguistically (RPG).

    The next paragraph is somewhat awkward.  [jwd]

    Why are modules not first-class? (RPG).  Abstraction to optimization is a
    poor link (RPG)---action: remove (JAP).  The ...recursive
    behaviour... definition is incorrect (RPG).  Implicit here is that {\tt
        symbol-value} and {\tt dynamic} are disjoint (RPG).

    JWD: We ought to consider whether we really want to require tail recursion
    optimization.  It is somewhat difficult to define precisely what the
    requirement means (see mail on the Scheme Standard list), and it makes it
    harder to use certain implementation strategies (eg, compile to C).  On the
    other hand, we don't have any other means for expressing iteration and I
    suspect many of us are inclined to follow Scheme.  A Rationale entry is
    required, at least.

    JWD: The Rationale should explain why dynamic variables are not by-module.
    It is possible to implement dynamic variables so that different vars with
    the same name can exist in different modules.  We would have to have some
    way to refer to (or import) variables in another module.

    Rewrote (once again) the description of the binding environments and gave
    them new names.  This removes the false distinction between lexical and
    module bindings that people complained about at Dec '90 meeting.  Any
    further complaints about this must be accompanied by revised wording.
\end{optPrivate}

\begin{optDefinition}
\eulisp\ is a dialect of Lisp and as such owes much to the great body of work
that has been done on language design in the name of Lisp over the last forty
years.  The distinguishing features of \eulisp\ are (i) the integration of the
classical Lisp type system and the object system into a single class hierarchy
(ii) the complementary abstraction facilities provided by the class and the
module mechanism (iii) support for concurrent execution.

Here is a brief summary of the main features of the language.
\begin{itemize}

    \item Classes\index{general}{class} are first-class objects.  The class
    structure integrates the primitive classes\index{general}{class!primitive}
    describing fundamental datatypes, the predefined classes and user-defined
    classes.

    \item Modules\index{general}{module} together with classes are the building
    blocks of both the \eulisp\ language and of applications written in \eulisp.
    The module system exists to limit access to objects by name.  That is,
    modules allow for hidden definitions.  Each module defines a fresh, empty,
    lexical environment\index{general}{module!environments}.

    \item Multiple control threads can be created in \eulisp\ and the
    concurrency model has been designed to allow consistency across a wide range
    of architectures.  Access to shared data can be controlled via locks
    (semaphores). Event-based programming is supported through a generic waiting
    function.

    \item Both functions and continuations are first-class in \eulisp, but the
    latter are not as general as in Scheme because they can only be used in the
    dynamic extent of their creation.  That implies they can only be used once.

    \item A condition mechanism which is fully integrated with both classes and
    threads, allows for the definition of generic handlers and supports both
    propagation of conditions and continuable handling.

    \item Dynamically scoped bindings\index{general}{binding!dynamically scoped}
    can be created in \eulisp, but their use is restricted, as in Scheme.
    \eulisp\ enforces a strong distinction between lexical bindings and dynamic
    bindings by requiring the manipulation of the latter via special forms.

\end{itemize}
%
\eulisp\ does not claim any particular Lisp dialect as its closest relative,
although parts of it were influenced by features found in
\cl\index{general}{cl@\cl}, \interlisp\index{general}{interlisp@\interlisp},
\lelisp\index{general}{lelisp@\lelisp},
\lisp/VM\index{general}{lispVM@\lisp/VM}, Scheme, and
T\index{general}{T}. \eulisp\ both introduces new ideas and takes from these
Lisps.  It also extends or simplifies their ideas as seen fit.  But this is not
the place for a detailed language comparison.  That can be drawn from the rest
of this text.

\eulisp\ breaks with \lisp\ tradition in describing all its types (in fact,
classes) in terms of an object system.  This is called The \eulisp\ Object
System, or \telos. \telos\ incorporates elements of the Common Lisp Object
System (CLOS)\index{general}{CLOS}~\bref{clos},
ObjVLisp\index{general}{ObjVLisp}~\bref{objvlisp},
Oaklisp\index{general}{Oaklisp}~\bref{oaklisp},
MicroCeyx\index{general}{MicroCeyx}~\bref{lelisp-manual}, and
MCS\index{general}{MCS}~\bref{mcs}.
%
\end{optDefinition}

\end{introduction}
