\defModule{string}{Strings}
%
\begin{optDefinition}
The defined name of this module is {\tt string}.  See also
section~\ref{collection} (collections) for further operations on
strings.
%
\syntaxform{string}
%
String literals\index{general}{external representation!string} are
delimited by the glyph called {\em quotation mark\/} (\verb+"+).  For
example, \verb+"abcd"+.

Sometimes it might be desirable to include string delimiter characters in
strings.  The aim of escaping in strings \index{general}{string!escaping in} is
to fulfill this need.  The \sc{string-escape}
\index{general}{string!string-escape glyph} glyph is defined as {\em reverse
    solidus\/} (\verb+\+).  String escaping can also be used to include certain
other characters that would otherwise be difficult to denote.  The set of named
special characters (see \S~\ref{character-set} and \S~\ref{character}) are
included in strings using the character digrams defined in
table~\ref{character-digrams}.  To allow arbitrary characters to appear in
strings, the hex-insertion digram is followed by an integer denoting the
position of the character in the current character set as for characters (see
\S~\ref{character-set}).  The syntax for the external representation of strings
is defined in syntax table~\ref{string-syntax} below:
%
\Syntax
\label{string-syntax}
\defSyntax{string}{
\begin{syntax}
    \scdef{string}: \\
    \>  " \scseqref{string-constituent} " \\
    \scdef{string-constituent}: \\
    \>  \scref{normal-string-constituent} \\
    \>  \scref{digram-string-constituent} \\
    \>  \scref{numeric string constituent} \\
   \scdef{normal-string-constituent}: \\
   \>  \scref{level-0-character} other than " or \textbackslash{} \\
   \scdef{digram-string-constituent}: one of \\
   \>  \textbackslash{}a \textbackslash{}b \textbackslash{}d \textbackslash{}f
   \textbackslash{}l \textbackslash{}n \textbackslash{}r \textbackslash{}t
   \textbackslash{}v \textbackslash{}" \textbackslash{}\textbackslash{} \\
   \scdef{numeric-string-constituent}: \\
   \>  \textbackslash{}x \scref{hexadecimal-digit} \\
   \>  \textbackslash{}x \scref{hexadecimal-digit} \scref{hexadecimal-digit} \\
   \>  \textbackslash{}x \scref{hexadecimal-digit} \scref{hexadecimal-digit} \\
   \>\>  \scref{hexadecimal-digit} \\
   \>  \textbackslash{}x \scref{hexadecimal-digit} \scref{hexadecimal-digit} \\
   \>\>  \scref{hexadecimal-digit} \scref{hexadecimal-digit}
\end{syntax}}%
\showSyntaxBox{string}

Some examples of string literals appear in table~\ref{example:string-literal}.
%
\begin{example}
\label{example:string-literal}
\examplecaption{Examples of string literals}
\begin{center}
\begin{tabular}{|ll|}\hline
    Example & Contents\\\hline
    \verb+"a\nb"+ & \verb+#\a+, \verb+#\n+ and \verb+#\b+\\
    \verb+"c\\"+ & \verb+#\c+ and \verb+#\\+\\
    \verb+"\x1 "+ & \verb+#\x1+ followed by \verb+#\space+\\
    \verb+"\xabcde"+ & \verb+#\xabcd+ followed by \verb+#\e+\\
    \verb+"\x1\x2"+ & \verb+#\x1+ followed by \verb+#\x2+\\
    \verb-"\x12+"- & \verb+#\x12+ followed by \verb-#\+-\\
    \verb+"\xabcg"+ & \verb+#\xabc+ followed by \verb+#\g+\\
    \verb+"\x00abc"+ & \verb+#\xab+ followed by \verb+#\c+\\\hline
\end{tabular}
\end{center}
\end{example}
%
\begin{note}
    At present this document refers to the ``current character set'' but
    defines no means of selecting alternative character sets.  This is to
    allow for future extensions and implementation-defined extensions
    which support more than one character set.
\end{note}
%
The function \functionref{write} outputs a re-readable form of any escaped
characters in the string.  For example, \verb+"a\n\\b"+ (input
notation) is the string containing the characters \verb+#\n+,
\verb+#\a+, \verb+#\\+ and \verb+#\b+.  The function \functionref{write}
produces \verb+"a\n\\b"+, whilst \functionref{prin} produces
%
\begin{verbatim}
a
\b
\end{verbatim}
%
The function \functionref{write} outputs characters which do not have a glyph
associated with their position in the character set as a hex insertion
in which all four hex digits are specified, even if there are leading
zeros, as in the last example in table~\ref{example:string-literal}.
The function \functionref{prin} outputs the interpretation of the characters
according to the definitions in section~\ref{character} without the
delimiting quotation marks.
%
\derivedclass{string}{character-sequence}
%
The class of all instances of \classref{string}.
%
\begin{initoptions}
%
\item[size, \classref{int}]
The number of characters in the string.  Strings are zero-based and
thus the maximum index is {\em size-1}.  If not specified the {\em
size\/} is zero.
%
\item[fill-value:, \classref{character}]
A character with which to initialize the string.  The default fill
character is \verb|#\x0|.
%
\end{initoptions}
%
\examples
%
\begin{tabular}{lcl}
\verb|(make <string>)| &\Ra& \verb|""|\\
\verb|(make <string> size: 2)| &\Ra& \verb|"\x0000\x0000"|\\
\verb|(make <string> size: 3| &\Ra& \verb|"aaa"|\\
\verb|  fill-value: #\a)|&&\\
\end{tabular}
%
\function{string?}
%
\begin{arguments}
    %
    \item[object] Object to examine.
    %
\end{arguments}
%
\result%
Returns {\em object\/} if it is a string, otherwise \nil{}.
%
\converter{symbol}
\begin{specargs}
    \item[string, \classref{string}] A string to be converted to a symbol.
\end{specargs}
%
\result%
If the result of \functionref{symbol-exists?} when applied to {\em string\/}
is a symbol, that symbol is returned.  If the result is \nil{}, then
a new symbol is constructed whose name is {\em string}.  This new
symbol is returned.
%
\method{binary=}{string}
%
\begin{specargs}
    \item[string$_1$, \classref{string}] A string.
    \item[string$_2$, \classref{string}] A string.
\end{specargs}
%
\result%
If the size of {\em string$_1$} is the same (under {\tt =}) as that of {\em
    string$_2$}, and the result of the conjunction of the pairwise application
of \methodref{binary=}{character} to the elements of the arguments is \true{}
the result is {\em string$_1$}.  If not the result is \nil{}.
%
\method{deep-copy}{string}
%
\begin{specargs}
    \item[string, \classref{string}] A string.
\end{specargs}
%
\result%
Constructs and returns a copy of {\em string\/} in which each element
is \functionref{eql} to the corresponding element in {\em string}.
%
\method{shallow-copy}{string}
%
\begin{specargs}
    \item[string, \classref{string}] A string.
\end{specargs}
%
\result%
Constructs and returns a copy of {\em string\/} in which each element
is \functionref{eql} to the corresponding element in {\em string}.
%
\method{binary<}{string}
%
\begin{specargs}
    \item[string$_1$, \classref{string}] A string.
    \item[string$_2$, \classref{string}] A string.
\end{specargs}
%
\result%
If the second argument is longer than the first, the result is \nil{}.
Otherwise, if the sequence of characters in {\em string$_1$} is pairwise less
than that in {\em string$_2$} according to \methodref{binary<}{character} the
result is \true.  Otherwise the result is \nil{}.  Since it is an error to
compare lower case, upper case and digit characters with any other kind than
themselves, so it is an error to compare two strings which require such
comparisons and the results are undefined.
%
\examples
\begin{tabular}{lcl}
\verb|(< "a" "b")| &\Ra& \verb|t|\\
\verb|(< "b" "a")| &\Ra& \verb|()|\\
\verb|(< "a" "a")| &\Ra& \verb|()|\\
\verb|(< "a" "ab")| &\Ra& \verb|t|\\
\verb|(< "ab" "a")| &\Ra& \verb|()|\\
\verb|(< "A" "B")| &\Ra& \verb|t|\\
\verb|(< "0" "1")| &\Ra& \verb|t|\\
\verb|(< "a1" "a2")| &\Ra& \verb|t|\\
\verb|(< "a1" "bb")| &\Ra& \verb|t|\\
\verb|(< "a1" "ab")| &\Ra& {\em undefined}
\end{tabular}
%
\seealso
Method \methodref{binary<}{character} (\ref{compare}) for characters (\ref{character}).
%
\method{as-lowercase}{string}
%
\begin{specargs}
    \item[string, \classref{string}] A string.
\end{specargs}
%
\result%
Returns a copy of {\em string\/} in which each character denoting an
upper case character, is replaced by a character denoting its lower
case counterpart.  The result must not be \functionref{eq} to {\em string}.
%
\method{as-uppercase}{string}
%
\begin{specargs}
    \item[string, \classref{string}] A string.
\end{specargs}
%
\result%
Returns a copy of {\em string\/} in which each character denoting an
lower case character, is replaced by a character denoting its upper
case counterpart.  The result must not be \functionref{eq} to {\em string}.
%
\method{generic-print}{string}
\begin{specargs}
    \item[string, \classref{string}] String to be ouptut on {\em stream}.
    \item[stream, \classref{stream}] Stream on which {\em string} is to be ouptut.
\end{specargs}
%
\result%
The string {\em string}.
%
Output external representation of {\em string\/} on {\em stream\/} as
described in the introduction to this section, interpreting each of
the characters in the string.  The opening and closing quotation marks
are not output.
%
\method{generic-write}{string}
\begin{specargs}
    \item[string, \classref{string}] String to be ouptut on {\em stream}.
    \item[stream, \classref{stream}] Stream on which {\em string\/} is to be ouptut.
\end{specargs}
%
\result%
The string {\em string}.
%
Output external representation of {\em string\/} on {\em stream\/} as
described in the introduction to this section, replacing single
characters with escape sequences if necessary.  Opening and closing
quotation marks are output.
\end{optDefinition}
